---
title: "1. Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(crayon.enabled = F)
```

```{r setup, warning=FALSE}
library(biPOD)
require(dplyr)
```

## Input format

The input for `biPOD` is a tibble with two columns:

- `time` denotes the instant of time in which each observation is collected
- `count` contains the values of the observations which should be a proxy of a population size (e.g. number of cells, volume, ...)

`biPOD` comes with a template dataset.

```{r}
# Load template data
data("xenografts", package = "biPOD")
```

## Initialisation of a new dataset

To use `biPOD` you need to initialize a `bipod` object with the initialisation function `init`.

When you create a new object it is required to explicit the sample name.

Additionally, since `biPOD` might deal with longitudinal observations coming from different time windows (e.g Natural evolution, Treatment, Relapse,...) divided by a set of `breakpoints`, the user should also input them if known. (see below)

## Example biPOD object

Let's use one of the sample coming from the `xenograft` data along with a made up set of breakpoints.

```{r}
mouse_id <- 529
x <- biPOD::init(
  counts = xenografts %>% dplyr::filter(mouse == mouse_id) %>% dplyr::mutate(count = tumour_volume),
  sample = mouse_id,
  break_points = c(0, 20)
)
```
If you look at the modified counts you will see that the observations have been grouped in different time windows

```{r}
biPOD::plot_input(x, add_highlights = T)
```
