---
title: "2. Task 1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{2. Task 1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=F}
library(biPOD)
require(dplyr)
```

The problem of characterizing a tumor growth model when growth rates change at known time-points while considering both exponential and logistic growths is denoted in `biPOD` as task 1. In this vignette we are going to look at how to use `biPOD` for this specific problem.

## Input data

We take the `xenografts` data and we make it ready for inference by preparing the `count` column and by dividing the `time` column by a factor of 7 (in order to work with a unit of time of a week).

```{r}
data("xenografts", package = "biPOD")
xenografts <- xenografts %>%
  dplyr::rename(count = tumour_volume) %>%
  dplyr::mutate(time = time / 7)
```

## Fit - Variational and no breakpoints

Let's take a sample which does not undergo treatment and should therefore exhibit a natural growth with a single parameter.
Hence, we are not going to consider any breakpoints for this sample.

```{r}
mouse_id <- 537
d <- xenografts %>%
  dplyr::filter(mouse == mouse_id)
x <- biPOD::init(
  counts = d,
  sample = mouse_id,
  break_points = NULL
)
```

Once the `bipod` object has been initialized it is possible to visualize the data using the `plot_input` function.

```{r}
biPOD::plot_input(x, log_scale = F, add_highlights = T)
```

Now, to fit the data for the first task one has to use the `fit` function. 

If you want to fit one specific type of growth (`exponential` or `logistic`) you have to pass one of the two to the parameter `growth_type`. Otherwise, if you want to test both of them and choose the best one you have to set the `growth_type` parameter to `both`.

Sometimes it might be helpful to scale the input data by a `factor_size`, so that all the counts will be divided by it.

When a population is decreasing or growing extremely slowly ($\rho \sim 0$) the instant of time in which the population was born might be impossible to infer. To deal with such situations, or when $t_0$ is of no interest, just set the `infer_t0` parameter to `FALSE`. 

In order to use a Variational Inference algorithm it suffice to set the `variational` parameter to `TRUE`.


```{r}
x <- biPOD::fit(
  x,
  growth_type = "both",
  model_selection_algo = "mixture_model",
  factor_size = 1,
  variational = T
)
```
## Plot - Variational and no breakpoints

### Diagnostics

The first plot to look at would be the one regarding the diagnostic of the inference. In this case, having used VI, we are going to look at the ELBO.
To do so, let's use the `plot_elbo` function which requires a `bipod` object and the specification of which elbo data we want to use since there might be multiple if we perform different task on the same object.

```{r, warning=F}
biPOD::plot_elbo(x, elbo_data = x$fit_elbo, diagnose = T)
```

### Plot of fit

Use the function `plot_fit` to plot the fit resulting after the inference of the first task.

Set `zoom` to `TRUE` to have a better view of the observations and `full_process` to `TRUE` to also plot the inferred value of $t_0$.

```{r, warning=F}
biPOD::plot_fit(x = x, zoom = T, full_process = T)
```

### Plot posteriors

You can use the general function `plot_posterior` to plot the posterior for a given parameters (use `plot_posteriors` for a set of parameters).

The names of the available parameters can be found in one of the fit objects inside the `bipod` object.
For example in this case let's plot $t_0$.

```{r, warning=F}
print(x$fit$summary() %>% dplyr::pull(variable))
biPOD::plot_posterior(
  x,
  x_fit = x$fit,
  par_name = "t0",
  color = "maroon"
)
```

### Model selection

Since we selected between exponential and logistic there is also the possibility to visualize this decision. In this case, having used the `mixture_model` algorithm we can plot it using the ``

```{r, warning=FALSE, message=F}
biPOD::plot_mixture_model_omega(x, plot_type = "boxplot", color = "forestgreen")
```

## Fit - MCMC and with breakpoints

Let's now consider a sample that undergoes treatment and should therefore exhibit a multiple growth rates. For this tutorial we will work with two manually chosen breakpoints (i.e 0 and 7).

```{r}
mouse_id <- 543
d <- xenografts %>%
  dplyr::filter(mouse == mouse_id)
x <- biPOD::init(
  counts = d,
  sample = mouse_id,
  break_points = c(0, 7)
)
```

Now the input will be grouped in different time windows, as it can also be seen by plotting it.

```{r, warning=F}
biPOD::plot_input(x, log_scale = T, add_highlights = T)
```

Let's use again the `fit` function. 

We want to use the MCMC algorithm so make sure to set the `variational` parameter to `FALSE`.

Moreover, we are going to use `bayes_factor` as a model selection algorithm.

```{r, warning=FALSE, message=F}
x <- biPOD::fit(
  x,
  growth_type = "both",
  model_selection_algo = "bayes_factor",
  variational = FALSE
)
```

## Plot - MCMC and with breakpoints

### Diagnostics

Let's plot the diagnostic of the inference. In this case, having used MCMC, we are going to look at the chains to see if they've mixed properly.
To do so, let's use the `plot_traces` function which requires a `bipod` object and the specification of which fit we want to use since there might be multiple if we perform different task on the same object.

```{r, warning=F}
biPOD::plot_traces(x, fit = x$fit, pars = NULL, diagnose = T)
```

### Plot of fit

Let's use the function `plot_fit`.
You can change the legend using the parameter `legend_labels` and `legend_title`.

```{r, warning=F, message=F}
biPOD::plot_fit(x = x, legend_labels = c("Natural growth", "Treatment", "Relapse"), legend_title = "Phase")
```

### Plot of posteriors

Let's use `plot_posteriors` function to plot all the growth rates one against each other.

```{r}
biPOD::plot_posteriors(
  x,
  x_fit = x$fit,
  par_list = c("rho[1]", "rho[2]", "rho[3]")
)
```

### Model selection

And finally, let's plot the Bayes factor using the `plot_bayes_factor` function.

```{r, warning=FALSE, message=F}
biPOD::plot_bayes_factor(x, with_categories = F)
```
