---
title: "3. Task 2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a3_task2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(biPOD)
require(dplyr)
```

In the previous task we assumed that the break points, i.e. the instant of time in which the latent dynamics changes abruptly, were know a priori. However, this is often not the case and one might be interested in inferring them.

## Input data

We take the `xenografts` data and we make it ready for inference by preparing the `count` column and by dividing the `time` column by a factor of 7 (in order to work with a unit of time of a week).

We select the sample 543, already presented in the previous vignette.

```{r}
data("xenografts", package = "biPOD")
mouse_id <- 543
d <- xenografts %>%
  dplyr::rename(count = tumour_volume) %>%
  dplyr::mutate(time = time / 7) %>%
  dplyr::filter(mouse == mouse_id)
```

## Breakpoints inference

You can infer the breakpoints in the following way. After having initialized a `bipod` object with no breakpoints you can use the function `breakpoints_inference`. 

If you reasonably know the possible number of breakpoints, you should pass a vector to the `available_changepoints`

```{r}
x <- biPOD::init(d, sample = mouse_id, break_points = NULL)
x <- biPOD::breakpoints_inference(x, available_changepoints = c(0,2), max_iter = 50)
```

The inferred breakpoints are now stored in the `metadata$breakpoints` field.

```{r}
print(x$metadata$breakpoints)
```

## Breakpoints posteriors

The posteriors of the inferred breakpoints can be seen using the `plot_breakpoints_posterior` function.

```{r}
biPOD::plot_breakpoints_posterior(x)
```

## Re-fit task 1

Now that the breakpoints have been inferred one can fit the data according to the first task simply using the `fit` function.
Let's see how it works on one example.

```{r}
x <- biPOD::fit(
  x,
  growth_type = "both"
)
```

And let's look at the final result.

```{r}
biPOD::plot_fit(x)
biPOD::plot_posteriors(
  x,
  x_fit = x$fit,
  par_list = c("rho[1]", "rho[2]", "rho[3]")
)
```
