---
title: "3. Inferring uknwokn breakpoints"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3. Inferring uknwokn breakpoints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(biPOD)
require(dplyr)
```

In the previous task we assumed that the break points, i.e. the instant of time in which the latent dynamics changes abruptly, were know a priori. However, this is often not the case and one might be interested in inferring them.

## Input data

We take the `xenografts` data and we make it ready for inference by preparing the `count` column and by dividing the `time` column by a factor of 7 (in order to work with a unit of time of a week).

We select the sample 543, already presented in the previous vignette.

```{r}
data("xenografts", package = "biPOD")
mouse_id <- 543
d <- xenografts %>%
  dplyr::rename(count = tumour_volume) %>%
  dplyr::mutate(time = time / 7) %>%
  dplyr::filter(mouse == mouse_id)
```

## Breakpoints inference

You can infer the breakpoints in the following way. After having initialized a `bipod` object with no breakpoints you can use the function `fit_breakpoints`. 

If you reasonably know the possible number of breakpoints, you should pass a vector to the `available_changepoints`.

```{r}
x <- biPOD::init(d, sample = mouse_id, break_points = NULL)
biPOD::plot_input(x)
x <- biPOD::fit_breakpoints(x, norm = T, n_trials = 500, avg_points_per_window = 3, available_changepoints = c(0,2), model_selection = "LOO", n_core=1)
x
```

The inferred breakpoints are now stored in the `metadata$breakpoints` field.

```{r}
print(x$metadata$breakpoints)
```

and their posterior distribution can be visualized using `plot_breakpoints_posterior` function


```{r}
biPOD::plot_breakpoints_posterior(x)
```

## Re-fit task 1

Now that the breakpoints have been inferred one can fit the data according to the first task simply using the `fit` function.
Let's see how it works on one example.

```{r}
x <- biPOD::fit(
  x,
  growth_type = "both",
  infer_t0 = F
)
```

And let's look at the final result.

```{r}
biPOD::plot_report(x, fit_type = 'simple')
```
