---
title: "4. Fitting a mixture of resistant-sensitive population"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{4. Fitting a mixture of resistant-sensitive population}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(biPOD)
require(dplyr)
```

When a population shrinks (e.g., sensitive to treatment) and later regrows (e.g., resistant to treatment) the dynamics are U-shaped. Here we are going to consider this specific case.

## Input data

We take the `xenografts` data and we make it ready for inference by preparing the `count` column and by dividing the `time` column by a factor of 7 (in order to work with a unit of time of a week).

We select the sample 543, already presented in the previous vignette and we filter out the observations that are prior to the treatment.

```{r}
data("xenografts", package = "biPOD")
mouse_id <- 543
d <- xenografts %>%
  dplyr::rename(count = tumour_volume) %>%
  dplyr::mutate(time = time / 7) %>%
  dplyr::filter(time >= 0) %>% 
  dplyr::filter(mouse == mouse_id)
x <- biPOD::init(d, "543 U-shape")
biPOD::plot_input(x)
```

## Two population model

The model assumes that

- only two populations exist, one dying out and one growing
- the growth rates are both positive and constant
- the sensible population must be present at x=0 but the same does not apply for the resistant one

and is able to infer

- $t_s$, instant of time in which the sensitive population dies out
- $t_r$, instant of time in which the resistant population was born
- $\rho_s$, death rate of the sensitive population
- $\rho_r$, growth rate of the resistant population

To do so, one need to use the `fit_two_pop_model` function.

```{r}
x <- biPOD::fit_two_pop_model(x, variational = F, factor_size = 1)
```

and the fit can be visualized as a single process

```{r}
biPOD::plot_two_pop_fit(x)
```

or splitting it depending on the two different populations

```{r}
biPOD::plot_two_pop_fit(x, split_process = T)
```
