---
title: "Get started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, warning = FALSE,           # hide messages/warnings by default
  fig.align = "center",                       # center all figures
  out.width = "80%",                          # shrink images a bit
  dpi = 200,                                  # crisper plots
  fig.retina = 2,                             # nicer on HiDPI screens
  dev = "png",
  echo = TRUE,
  collapse = TRUE,                            # console-style output
  comment = "#>"                              # prefix for outputs
)
```

```{r}
#| message: false
#| warning: false
# install.packages("devtools")
# devtools::install_github("caravagnalab/biPOD")
library(biPOD)
require(dplyr)
set.seed(1)
```

## What is biPOD?

biPOD is an R package for **model-based Bayesian inference of population dynamics** from heterogeneous longitudinal data.
It supports: (i) fitting growth with **known breakpoints**, (ii) **inferring breakpoints** when unknown, and (iii) a **two-population (sensitive/resistant)** model typical in treatment–relapse settings.

## 1) Quick tour with the built-in dataset

The package ships the `xenografts` mouse model dataset:

```r
data(xenografts)
head(xenografts)
```

We initialize a `bipod` object and plot inputs:

```{r}
mouse_id = 544
df = xenografts %>% 
  dplyr::filter(mouse == mouse_id) %>% 
  dplyr::mutate(count = tumour_volume)
x <- init(counts = df, sample = unique(df$mouse), break_points = NULL)
plot_input(x)
```


### Fit with known breakpoints

Suppose change-points (e.g., treatment start and end) are known:

```{r}
x <- biPOD::init(
  counts = df,
  sample = mouse_id,
  break_points = c(0, 50)
)
x <- biPOD::fit(
  x = x, 
  growth_type = "exponential", 
  infer_t0 = T
)
plot_fit(x)
plot_posteriors(x = x, x_fit = x$fit, par_list = c("rho[1]", "rho[2]", "rho[3]"))     # posterior densities for parameters
```

### Infer breakpoints

When transition times are **unknown**, estimate them directly from the data:

```{r}
x <- biPOD::init(
  counts = df,
  sample = mouse_id,
  break_points = NULL
)
x <- biPOD::fit_breakpoints(x = x, available_changepoints = 0:2)
plot_breakpoints_posterior(x) + ggplot2::geom_vline(xintercept = c(0, 50))
```


## Session info

```r
sessionInfo()
```
