---
title: "4. Uncertainty and number of points"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(biPOD)
```

The uncertainty over the parameters inferred by the model obviously depends on the number of points available to fit the model. 
We will simulate a process lasting 1 unit of time and with 10 total time steps. We will then fit a model for a number of observation starting from 1 up to 10.
Initially we will observe only the starting and the final population, but we will then iteratively add a random new observation to the sample.

This is the full simulated process:

```{r}
n_steps <- 10
dt <- .1
d <- biPOD::sim_stochastic_exponential(20, 1, .25, steps = n_steps, delta_t = dt)
x <- biPOD::init(d, "Full sample")
biPOD::plot_input(x)
```

```{r, results=FALSE, message=FALSE}
indexes <- c(1, 11)
indexes <- c(indexes, sample(2:n_steps, n_steps - 1, replace=F))

plot_fits <- list()
posterior_plots <- list()

for (i in 3:length(indexes)) {
  idx <- sort(indexes[1:i])
  
  x <- biPOD::init(d[idx,], sample = paste(""))
  x <- biPOD::fit(x, growth_type = "exponential", t0_lower_bound = -20, factor_size = .01)
  
  name = paste("sample", i)
  print(name)
  print(d[idx,])
  
  plot_fits[[name]] = biPOD::plot_simple_fit(x, full_process = F)
  posterior_plots[[name]] = ggpubr::ggarrange(plotlist = biPOD::get_growth_rate_posteriors(x))
}
```

```{r}
ggpubr::ggarrange(plotlist = plot_fits)
ggpubr::ggarrange(plotlist = posterior_plots)
```
